\chapter{Dynamic IFC with Unstructured Control Flow and Exceptions}
\label{ch:ucfe}
\blfootnote{The content of this
  chapter is based partly on the work published as part of the paper,
  ``Information Flow Control in WebKit's JavaScript
  Bytecode''~\cite{post14}} 
This chapter presents a mechanism to prevent leaks due to implicit
flows in the presence of unstructured control constructs like break,
return-in-the-middle and exceptions.  

Implicit flow corresponds to \emph{control dependence} in program
analysis, where a predicate governs which program path is executed and
leaks information through the control flow of the program. To avoid
overtainting $\pc$ labels, an important goal in implicit flow 
tracking is to determine when the influence of a control construct has
ended. For block-structured control flow limited to \texttt{if} and
\texttt{while} commands, this is straightforward: The effect of a
control construct ends with its lexical scope, e.g., in 
$$\texttt{if (h) \{l = 1;\} l = 2}$$
\texttt{h} influences the control flow at
\texttt{l = 1} but not at \texttt{l = 2}. This leads to a
straightforward implementation of a $\pc$ upgrading and downgrading
strategy: One maintains a \emph{stack} of $\pc$
labels~\cite{zdancewic02PhD}; the effective $\pc$ is the top one. When
entering a control flow construct like \texttt{if} or \texttt{while},
a new $\pc$ label, equal to the join of labels of all values on which
the construct's guard depends with the previous effective $\pc$, is
pushed. When exiting the construct, the label is popped.

Unfortunately, it is unclear how to extend this simple strategy to
non-block-struct\-ured control flow constructs such as 
\texttt{break}, \texttt{continue} and \texttt{return}-in-the-middle
for functions, all of which occur in high-level languages. For
example, consider the program 
$$\texttt{l = 1; while(1) \{... if(h) \{break;\}; l = 0; break;\}}$$ 
with \texttt{h} labeled $H$. This program leaks the
value of \texttt{h} into \texttt{l}, but no assignment to \texttt{l}
appears in a block-scope guarded by \texttt{h}. Indeed, the $\pc$
upgrading and downgrading strategy just described is ineffective for
this program. 

Implicit flow in the form of error handling is also a  
source of information leak as it helps the adversary to learn about
the system~\cite{leak}. For instance, an exception handler might print 
the stack trace of the error, from which an attacker can determine what
sort of attacks the system is vulnerable to. 

Tracking information flow in the presence of unstructured control
flows is non-trivial as the control breaks out of block
structures. Exceptions are much more difficult to handle as they 
allow for non-local control transfer. Much work on error handling has
focused in the context of static analysis~\cite{myersJFlow,aslan_plas09}
and the work on IFC for dynamic languages has mostly ignored
exceptions and other unstructured control flow 
constructs~\cite{plas09,plas10,stagedIFC,guarnieri11ISSTA,acsac09}.
Just et al.~\cite{just11PLASTIC} present dynamic IFC for JavaScript
bytecode with static analysis to determine implicit flows precisely
but ignore implicit flows due to exceptions. Hedin and Sabelfeld
propose a dynamic IFC approach for a language 
modeling the core features of JavaScript~\cite{csf12} but ignore
unstructured control flow constructs like \emph{break},
\emph{continue} and \emph{return}-in-the-middle for functions. For
handling exceptions, they introduce annotations and an additional
class of labels. An extension introduces similar annotations to deal
with unstructured control flows~\cite{sac14}. These labels are more
restrictive than needed, e.g., the code indicated by dots in the
example above is executed irrespective of the condition \texttt{h} in
the first iteration, and thus there is no need to raise the $\pc$
before checking that condition. 

To solve this issue, this chapter presents a precise dynamic
analysis approach using \emph{post-dominator}
analysis~\cite{denning82,just11PLASTIC}.  

\section{Control Flow Graphs and Post-dominator Analysis}
The approach presented in this chapter performs on-the-fly
post-dominator analysis at runtime to handle implicit
flows. A control flow graph (CFG), which is a directed graph, is
constructed for every new function before it is executed with every
instruction being represented as a node and whose edges represent the
possible control flows. For every branch node, its immediate
post-dominator (IPD) is
computed~\cite{denning82,just11PLASTIC,post14}. A stack of $\pc$  
labels is maintained. When executing a branch node, a new $pc$ label
is pushed on the stack \emph{along with} the node's IPD. When the IPD
is actually reached, the $pc$ label along with the IPD is popped from
the stack. In~\cite{Xin,Masri}, the authors prove that the IPD marks the
end of the scope of an operation and hence the security context of the
operation, so our strategy is sound. The IPD-based solution works for all
forms of unstructured control flow like \emph{break}, \emph{continue},
\emph{return}-in-the-middle, and exceptions. Multiple \emph{return}
statements in a function are represented by a single \emph{return} 
node. Theorem~\ref{thm1:exc} shows that the IPD of a node is the most
precise node where the context of an operation can be removed. For
proving Theorem~\ref{thm1:exc}, a few definitions are defined below.

\begin{mydef}\label{defcfg}
\emph{(Control flow graph)}\\
A \emph{control flow graph} is a directed graph $\cfg =
(\mathcal{N}, \mathcal{E}, n_s, n_e, \mathcal{L})$. $\mathcal{N}$ is the set of
nodes. $\mathcal{E}$ is the set of control flow edges $(n_1, n_2)
\in \mathcal{E}$, where $n_i \in \mathcal{N}$. $(n_1, n_2)$ represents $n_2$ may
immediately execute after $n_1$. The nodes $n_s, n_e \in \mathcal{N}$
are special nodes representing the start and end point of $\cfg$.  The
function $\mathcal{L}$ maps the edges in $\mathcal{E}$ to labels.
\end{mydef}

\begin{mydef}\label{defpath}
\emph{(Path)}\\
A \emph{path} in a CFG $\cfg$ is a sequence of nodes $(n_1, n_2, ...,
n_m)$ such that $(n_i, n_{i+1}) \in \mathcal{E}$, written as
$\pathG{n_1}{n_m}$. 
A node $n$ that lies on the path $\pathG{n_1}{n_m}$ is written as
$n \in \pathG{n_1}{n_m}$.
The notation $n_1 < n_2$ with respect to two nodes $n_1$ and $n_2$ in
a CFG $\cfg$ indicates that $n_2$ lies on a path $\pathG{n_1}{n_e}$ . 
\end{mydef}

\begin{mydef}\label{defbp}
\emph{(Branch-point)}\\
A \emph{branch-point} $b$ is a node in a CFG $\cfg$ that has more than
one successor, i.e., \\ outdegree$(b) > 1$.
\end{mydef}

\begin{mydef}\label{defpd}
\emph{(Post-dominator)}\\
In a CFG $\cfg$, a node $n_d$ is said to be the \emph{post-dominator} of a node
$n$ if all paths from $n$ to the end-node pass through $n_d$, i.e.,
$\forall p. \pathG{n}{n_e} \implies n_d \in p$. The notation
\emph{$\dom{n_d}{n}$} indicates that $n_d$ is a post-dominator of $n$.
\end{mydef}

\begin{mydef}\label{defipd}
\emph{(Immediate post-dominator)}\\
A node $i$ is the immediate post-dominator of the node $n$, denoted as
$\IPD{n}$, if{f}:
\begin{enumerate}
\item \emph{$\dom{i}{n}$} and
\item \emph{$\not\exists n_o \in \mathcal{N}.((\dom{n_o}{n}) \wedge (n_o <
  i))$} or \\
  \emph{$\forall n_o \in \mathcal{N}.((n_o \neq n) \implies (\dom{n_o}{n}
  \implies \dom{n_o}{i}))$}.
\end{enumerate}
\end{mydef}

\begin{myThm}[Precision]\label{thm1:exc}
Choosing any node other than the IPD to lower the $\pc$-label will either
give unsound results or be less precise.
\end{myThm}
\begin{proof}
The proof of the theorem is described in Appendix~\ref{app:exc-pre}.
\end{proof}

\section{Exceptions and Synthetic Exit Nodes}
\label{sec:excsen}
Maintaining a precise CFG for post-dominator analysis at runtime in the 
presence of exceptions is expensive. The CFG of a function is 
constructed statically on-the-fly when compiling the function at 
runtime. An exception-throwing node in a
function that does not catch that exception should have an outgoing
control flow edge to the next related exception handler in the runtime
call-stack. This means that the CFG is, in general, inter-procedural,
and edges going out of a function depend on its calling context, so
IPDs of nodes in the function must be computed \emph{every time the
  function is called} (the IPDs change based on the earlier functions
in the call-stack that called the particular function where the
exception occurs). Moreover, in the case of recursive functions, the
nodes must be replicated for every call. This is rather
expensive. Ideally, the function's CFG should be built only once when
\emph{the function is compiled} and work intra-procedurally. 

In the design presented in this chapter, every function that may throw
an unhandled exception and has an exception handler present earlier in
the call-stack (which is assumed to be known at runtime through 
additional data structures in the system) has a special
\emph{synthetic exit node} ($\SEN$), which is placed after the regular
return node of the function in the CFG. 
Every exception-throwing node, whose exception will not be caught
within the function, has an outgoing edge to the $\SEN$. In essence, the
$\SEN$ is treated as the IPD for nodes whose actual IPDs lie outside of
the function. By doing this, all cross-function edges are eliminated
and the CFGs become intra-procedural. This allows the computation of
the CFGs just once as compared to the inter-procedural case. 
% As there are instructions other than \emph{throw} can also throw exceptions, we
% treat them as exception-throwing instructions and calculate the IPDs
% accordingly, i.e., 
For every exc\-ep\-tion-throw\-ing instruction that has an associated
handler, its context is maintained during dynamic information flow
analysis until the handler is reached. 
% This is a bit restrictive; hence, this option can be turned
% off in the instrumented interpreter, if required. 
Thus, function calls and all potential exception-throwing
instructions are represented as nodes with multiple edges (bra\-nches) 
and push a node on the $\pc$-stack. However, a new node is \emph{not} 
pushed on the $\pc$-stack if the IPD of the current node is the same
as the IPD on the top of the $\pc$-stack or if the IPD of the current
node is the $\SEN$, as in this case the \emph{real} IPD, which is outside
of this method, is already on top of the $\pc$-stack. In fact, the
actual IPD of a node having $\SEN$ as its IPD is the node that is
currently on the top of the stack. This result is shown in
Theorem~\ref{thm2:exc}. The proof is shown in Appendix~\ref{app:exc-pre}.

\begin{myThm}\label{thm2:exc}
The actual IPD of a node having $\SEN$ as its IPD is the node on
the top of the pc-stack, which lies in a previously called function.
\end{myThm}

In summary, these semantics emulate the effect of having cross-function edges. 
For illustration, consider the following two functions \texttt{f} and
\texttt{g}. The $\diamond$ at the end of \texttt{g} denotes its
$\SEN$. Note that there is an edge from \texttt{throw 9} to $\diamond$
because \texttt{throw 9} is not handled within~\texttt{g}. $\Box$
denotes the IPD of the function call~\texttt{g()} and handler~\texttt{catch(e)}.

%\noindent 
%\begin{tabular}{@{}p{0.5\textwidth}p{0.4\textwidth}@{}}
%\begin{lstlisting}[mathescape=true]
%function f() = {
%  l = 0;
%  try { g(); } catch(e) { l = 1; }
%  $\Box$ return l;
%}
%\end{lstlisting} & 
%\begin{lstlisting}[mathescape=true]
%function g() = {
%  if (h) {throw 9;}
%  return 7;
%} $\diamond$
%\end{lstlisting}
%\end{tabular}

\noindent 
\begin{tabular}{@{}p{0.55\textwidth}p{0.35\textwidth}@{}}
\begin{tabular}[t]{@{}l@{}}
\texttt{function f() = \{} \\
~~\texttt{l = 0;} \\
~~\texttt{try \{ g(); \} catch(e) \{ l = 1; \}} \\
~~$\Box$ \texttt{return l;}\\
\texttt{\}}
\end{tabular} & 
\begin{tabular}[t]{@{}l@{}}
\texttt{function g() = \{}\\
~~\texttt{if (h) \{throw 9;\}}\\
~~\texttt{return 7;}\\
\texttt{\}} $\diamond$
\end{tabular}
\end{tabular} 

In the absence of instrumentation, when
\texttt{f} is invoked with $\pc = L$, the two functions together leak
the value of \texttt{h}, which is assumed to have a label $H$, into the
return value of \texttt{f}. When calling \texttt{g}, the current $\pc$
and IPD $(L, \Box)$ are pushed on the $\pc$-stack.  When executing the
condition \texttt{if (h)} a new node is not pushed again, but the top
element is merely updated to $(H, \Box)$ as its IPD is the $\SEN$
$\diamond$. If \texttt{h} is \texttt{false}, control reaches the
\texttt{return} statement but with $\pc = H$.
% because the IPD of \texttt{if (h)} is $\diamond$ and the real IPD is $\Box$. 
At $\Box$, $\pc$ is lowered to $L$, so
\texttt{f} ends with the return value \texttt{0} and public label
$L$. If \texttt{h} is \texttt{true}, control reaches the handler, which
is in \texttt{f} and invokes it with the same $\pc$ as at the point of
exception, i.e., $H$. Consequently, permissive-upgrade marks the 
assignment in the catch block as partially-leaked and prevents the
implicit information leak in this case. 

\input{chapters/ucf/formal}