\chapter{Information Flow Policies for Web Browsers}
\label{ch:webpol}
\blfootnote{The content of this chapter is based on the work published
  as part of the paper,  ``WebPol: Fine-grained Information Flow
  Policies for Web Browsers''~\cite{webpol}}  

% Existing web security standards and web browsers address the problem
% of information leaks unsatisfactorily, favoring functionality over
% privacy. Most of these mechanisms are primarily access control
% mechanisms describing whether or not a script can access certain data,
% whereas the meaningful requirement is to allow untrusted scripts
% access to confidential data that they need and to prevent the scripts
% from leaking the data on the side. Recently, a number of solutions
% based on information flow control (IFC) have been
% proposed~\cite{flowfox,csf14,jsflow,post14,csf15,chudnov-ccs,cowl,jang10CCS}
% to allow third-party scripts access to necessary sensitive data, but
% restrict \emph{where} scripts can send the data (and its derivatives)
% in accordance with a security policy. While this would balance
% functionality and privacy perfectly, all existing IFC-based solutions
% for web browsers fall short of this ideal goal.
Many approaches~\cite{jsflow,post14,csf15,chudnov-ccs}, have been
proposed for enforcing dynamic information flow control in web
browsers, but currently lack adequate support for specifying policies
conveniently. Flowfox~\cite{csf14} provides a rich policy framework
but all websites are subject to the same policy, and the underlying IFC
technique, secure multi-execution~\cite{SME}, does not handle shared
state soundly. COWL~\cite{cowl} uses coarse-grained isolation,
allowing scripts' access to either remote domains or the shared state,
but not both. This requires significant code changes when both are
needed simultaneously (see Chapter~\ref{ch:related} for more
details).

This chapter presents {\sys}, a policy framework that allows
a webpage developer to release data selectively to third-party scripts
(to obtain useful functionality), yet control what the scripts can do
with the data. {\sys} policies label
sensitive content (page elements and user-generated events) at source,
and selectively declassify them by specifying where (to which domains)
the content and its derivatives can flow. Host page developers specify
{\sys} policies in JavaScript, a language already familiar to them. 
{\sys} is integrated with the instrumentation of the dynamic IFC
framework for WebKit presented thus far ({\sys} integrates with any
taint-based IFC solution to overcome the shortcomings listed
above). The expressiveness of {\sys} policies is demonstrated through
examples and by applying {\sys} to two real websites. 

\input{chapters/browser/webpol/overview}
\input{chapters/browser/webpol/model}
\input{chapters/browser/webpol/examples}
