\section*{Acknowledgment}

My first thanks goes to my primary advisor, Christian Hammer, to believe 
in me and offer me a research position. Apart from teaching me
important lessons in research and shaping my development as a PhD
student, he helped me get settled at a new home away from home. He has
been highly supportive in whatever initiatives I took. Additionally, I
would like to thank my secondary advisor, Deepak Garg, who taught me
about how foundations are necessary for building a proper
system. Besides, the other not-so-research based talks and advices
from both have helped me take further steps in my career. The
acknowledgment would be incomplete without mentioning my master's
supervisor, Prof. Ramesh Joshi, who introduced me to the wonderful
field of research at IIT Roorkee.

Naturally, many thanks also go to my closest collaborator and
co-author, Vineet Rajani, without whom this thesis would not have been
possible and with whom I had really great and fruitful discussions on
security. Together, we had a lot of fun while doing research and
outside of it. I would also like to thank my friends and colleagues
--- Abhishek Tiwari, Aravinda Krishnan, Himangshu Saikia, and
Kaustubh Beedkar, who added loads of fun and food to my
PhD. In particular, I want to mention my colleagues
David Pfaff, Jenny Hotzkow, and Marie-Therese Walter, with whom I had
the pleasure of sharing the office and who helped me in handling all
the German documents and procedures, and also writing the German
abstract of my dissertation. Special thanks to CISPA
secretaries Stephanie Feyahn, Isa Maurer, and Sabine Nermerich, who
literally did all the bureaucratic tasks that I had to and helped me
in every way possible.

I would also like to acknowledge Karthikeyan
Bhargavan, Aseem Rastogi and Kapil Vaswani to have given me the
opportunity to work as a research intern with them. It was a wonderful 
and highly interesting work environment for me and greatly helped
me explore other related fields of research. I met and made friends
with Rahul Kidambi, Sameer Wagh, and Shruti Tople during these
internships who also helped me in making career choices besides
all the fun-filled game and movie sessions. 

I am deeply grateful to my family and friends
for their unconditional love, their faith in me, and their advice in all
situations. Without them, I would not have been able to go my own way
during the years of my PhD and would certainly not stand where I am 
today. 

